---
title: "HW2"
author: "Hannah Boyd"
date: "`r Sys.Date()`"
output:
  html_document: default
---
```{r}
library(rafalib)
library(dplyr)
library(downloader)
url<-"https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/mice_pheno.csv"
filename<-basename(url)
download(url,destfile=filename)
dat<-na.omit(read.csv(filename))
```

# Question 1

## Part a

```{r}
pnorm(1)-pnorm(-1)
```

**Answer: 0.6826895 are within one standard deviation**

## Part b

```{r}
pnorm(2)-pnorm(-2)
```

**Answer: 0.9544997 are within two standard deviations**

## Part c

```{r}
pnorm(3)-pnorm(-3)
```

**Answer: 0.9973002 are within three standard deviations**

# Question 2

```{r}
y <- dat %>% filter(Sex == 'M' & Diet == 'chow')%>% dplyr::select(Bodyweight) %>% unlist()
mean(y) 
popsd(y)
z_score <- (y - mean(y))/popsd(y)

```

## Part a

```{r}
mean(abs(z_score)<= 1)
```

**Answer: 0.6950673 are the proportion of mice within one SD from the average weight**

## Part b

```{r}
mean(abs(z_score)<= 2)
```

**Answer: 0.9461883 are the proportion of mice within two SD from the average weight**

## Part c

```{r}
mean(abs(z_score)<= 3)
```

**Answer: 0.9910314 are the proportion of mice within three SD from the average weight**

# Question 3

```{r}
mypar(1,1)
qqnorm(y)
qqline(y)
```

**Answer: C.) The mouse weights are well approximated by the normal distribution, although the larger values (right tail) are larger than predicted by the normal** 

# Question 4

## Part a

```{r}
y <- filter(dat, Sex=="M" & Diet=="chow") %>% dplyr::select(Bodyweight) %>% unlist
avgs <- replicate(10000, mean( sample(y, 25)))
mypar(1,2)
hist(avgs)
qqnorm(avgs)
qqline(avgs)
```

## Part b 
```{r}
m <- 10000
n <- 25
y <- filter(dat, Sex=='M' & Diet=='chow') %>%dplyr::select(Bodyweight) %>% unlist() 
set.seed(1)
avg_list <- replicate(m,{ + mean(sample(y,25))})
mypar(1,2)
hist(avg_list)
qqnorm(avg_list)
qqline(avg_list)
mean(avg_list)
```

**The average of the distribution of the sample average is 30.96856.**

## Part c

```{r}
popsd(avg_list)
```

**Answer: The standard deviation is 0.827082**

## Part d

```{r}
popsd(y)/sqrt(25)
```

**Answer: D**

# Question 5

```{r}
m <- 10000
set.seed(1)
sd_list <- replicate(m,{sd(sample(y,25))})
mypar(1,1)
hist(sd_list)
mean(sd_list <= 3.5)
```

**Answer: The proportion below 3.5 is 0.0942.**

# Question 6

```{r}
x = seq(0.0001, 0.9999, len = 300)
df_list <- c(3,10,30,100)
mypar(2,2)
for (i in seq_along(df_list)) {
qqnorm(qt(x,df_list[i]), main = df_list[i])}
```

**Answer: C** 


