---
title: "HW3"
author: "Hannah Boyd"
date: "`r Sys.Date()`"
output: html_document
---

# Question 1

```{r}
url<-"https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/babies.txt"
library(utils)
library(dplyr)
filename<- basename(url)
download.file(url,destfile=filename)
babies<-read.table("babies.txt",header=TRUE)
bwt.nonsmoke<-filter(babies,smoke==0)%>%select(bwt)%>%unlist
bwt.smoke<-filter(babies,smoke==1)%>%select(bwt)%>%unlist
```

## Part 1

```{r}
N = 10
set.seed(1)
nonsmokers <- sample(bwt.nonsmoke , N)
smokers <- sample(bwt.smoke , N)
obs <- mean(smokers) - mean(nonsmokers)
```

```{r}
dat <- c(smokers,nonsmokers)
shuffle <- sample( dat )
smokersstar <- shuffle[1:N]
nonsmokersstar <- shuffle[(N+1):(2*N)]
mean(smokersstar)- mean(nonsmokersstar)
```
```{r}
obs <- mean(smokers) - mean(nonsmokers)

FUN <- function(dat){
     N <-  length(dat) / 2
     dat <- c(smokers,nonsmokers)
     shuffle <- sample( dat )
     smokersstar <- shuffle[1:N]
     nonsmokersstar <- shuffle[(N+1):(2*N)]
     mean(smokersstar)-mean(nonsmokersstar)
 }

set.seed(1)
permuted <- replicate(999,FUN(dat))
dist <- append(permuted,obs)

p.value <- sum(dist <= obs) / 1000
p.value
```
**Answer: The p value for a 1,000 permutations of the mean is 0.051.** 

## Part II

```{r}
obs <- median(smokers) - median(nonsmokers)
    
FUN <- function(dat){
     N <-  length(dat) / 2
     dat <- c(smokers,nonsmokers)
     shuffle <- sample( dat )
     smokersstar <- shuffle[1:N]
     nonsmokersstar <- shuffle[(N+1):(2*N)]
     median(smokersstar)-median(nonsmokersstar)
 }

set.seed(1)
permuted <- replicate(999,FUN(dat))
dist <- append(permuted,obs)
p.value <- sum(dist <= obs) / 1000
p.value
```

**Answer: The p value for 1,000 permutation of the median is 0.128**

# Question 2 

```{r}
library(datasets)
data("ChickWeight")
head(ChickWeight)
plot(ChickWeight$Time, ChickWeight$weight, col=ChickWeight$Diet)
chick = reshape(ChickWeight, idvar=c("Chick","Diet"), timevar="Time",direction="wide")
chick = na.omit(chick)
chick$weight.4
```

## Part 1A

```{r}
day4 <- chick$weight.4
day4plus <- c(day4,3000)
mean(chick$weight.4)
mean(day4plus)
mean(day4plus)-mean(chick$weight.4)
```

**Answer:The mean increases by 63.90966 with the addition of the outlier.**

## Part 1B

```{r}
median(day4plus)
median(chick$weight.4)
median(day4plus)-median(chick$weight.4)
```

**Answer: The median does not increase with the addition of the outlier.**

## Part1C

```{r}
sd(day4plus)
sd(chick$weight.4)
sd(day4plus)-sd(chick$weight.4)
```

**Answer: the standard deviation increases by 429.1973 with the addition of the outlier.**

## Part1D

```{r}
mad(day4plus)
mad(chick$weight.4)
mad(day4plus)-mad(chick$weight.4)
```

**Answer: the median absolute deviation does not increase with the addition of the outlier**

## Part 2A

```{r}
x <- chick[chick$Diet==1,"weight.4"]
y <- chick[chick$Diet==4,"weight.4"]
t.test(x,y)
```

```{r}
wilcox.test(x,y)$p.value
```

## Part 3

```{r}
set.seed(38011)
avgdiff <- replicate(1000, {
    dat <- c(x,y)
    shuffle <- sample(dat) 
    xstar <- shuffle[1:N] 
    ystar <- shuffle[(N+1):(2*N)]
    return(mean(xstar)-mean(ystar))})
 
(sum(abs(avgdiff) > abs(obs)) + 1) / (length(avgdiff) + 1)
```


**Answer: The permutation pvalue is larger than the p value for the t test (7.32e-06) or the wilcox test (0.0002011939)**

## Part 4

```{r}
yplus<-c(y,1000)
t.test(x,yplus)
```

**Answer: The outlier increased the p value for both the t test and the wilcox test.**


## Part 5

```{r}
set.seed(58381)
avgdiff <- replicate(1000, {
dat <- c(x,y)
shuffle <- sample(dat) 
xstar <- shuffle[1:N] 
ystar <- shuffle[(N+1):(2*N)]
return(mean(xstar)-mean(ystar))})
(sum(abs(avgdiff) > abs(obs)) + 1) / (length(avgdiff) + 1)

```

**Answer: The p value did not change despite the set.seed value changing.The permutation pvalue is larger than the p value for the t test (7.32e-06) or the wilcox test (0.0002011939) from the dataset without the outlier.**
