---
title: "HW4"
author: "Hannah Boyd"
date: "`r Sys.Date()`"
output: html_document
---

# Question 1

```{r}
library(devtools)
install_github("genomicsclass/GSE5859Subset")
library(GSE5859Subset)
data(GSE5859Subset)
```
```{r}
library(genefilter)
?rowttests
```

```{r}
sum(rowttests(geneExpression, as.factor(sampleInfo$group))$p.value < 0.05)
```
**Answer: There are 1383 genes with p values smaller than 0.05.**

# Question 2

```{r}
sum(rowttests(geneExpression, as.factor(sampleInfo$group))$p.value < (0.05/nrow(geneExpression))) 
```
**Answer: With the bonferroni corrections, only 10 p values are still significant.**

# Question 3 

```{r}
pvals = rowttests(geneExpression, as.factor(sampleInfo$group))$p.value
sum(p.adjust(pvals, method = "fdr", n = length(pvals))<0.05)
```
**Answer: Using the p.adjust function, 13 genes have a p value less than 0.05.**

# Question 4 

```{r}
library(qvalue)
sum(qvalue(pvals)$qvalues<0.05)
```
**Answer: There are 22 genes with q values less than 0.05**

# Question 5

```{r}
qvalue(pvals)$pi0
```
**Answer: The estimated proportion of genes for which the null hypothesis is true is 0.6695739** 

# Question 6

```{r}
plot(qvalue(pvals)$qvalue/p.adjust(pvals,method="fdr"))
abline(h=qvalue(pvals)$pi0,col=2)
```


**Answer: C.) the qvalue function estimates the proportion of genes for which the null hypothesis is true and provides a less conservative estimate**

# Question 7

```{r}
n <- 24
m <- 8793
mat <- matrix(rnorm(n*m),m,n)
```

```{r}
delta <- 1
positives <- 500
mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
```

## Part I - False Positives

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
m0=8293
B <- 1000
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
   FP1 <- sum(pvals[-(1:positives)]<=0.05/m)  
  })
mean(result/m0)
```

### p.adjust

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
m0=8293
m1=500
B <- 1000
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
  padj = p.adjust(pvals, method = "fdr", n = m)
  FP1 <- sum(padj[-(1:positives)]<=0.05)
  })
mean(result/m0)
```

### qvalue

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
m0=8293
m1=500
B <- 1000
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
  padj = qvalue(pvals)$qvalues
  FP1 <- sum(padj[-(1:positives)]<=0.05)
  FN1 <- sum(padj[1:positives]>0.05)
  c(FP1,FN1)
  })
mean(result[1,]/m0)
```


## Part II - False Negatives

### Bonferroni

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
B <- 1000
m0=8293
m1=500
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
  ##Bonferroni
  FP1 <- sum(pvals[-(1:positives)]<=0.05/m)  
  FN1 <- sum(pvals[1:positives]>0.05/m)
  c(FP1,FN1)
})
mean(result[2,]/m1)
```

### p.adjust

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
m0=8293
m1=500
B <- 1000
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
  padj = p.adjust(pvals, method = "fdr", n = m)
  FP1 <- sum(padj[-(1:positives)]<=0.05)
  FN1 <- sum(padj[1:positives]>0.05)
  c(FP1,FN1)
  })
mean(result[2,]/m1)
```


### qvalue 

```{r}
set.seed(1)
library(qvalue)
library(genefilter)
n <- 24
m <- 8793
m0=8293
m1=500
B <- 1000
delta <-2
positives <- 500
g <- factor(rep(c(0,1),each=12))
result <- replicate(B,{
  mat <- matrix(rnorm(n*m),m,n)
  mat[1:positives,1:(n/2)] <- mat[1:positives,1:(n/2)]+delta
  pvals = rowttests(mat,g)$p.val
  padj = qvalue(pvals)$qvalues
  FP1 <- sum(padj[-(1:positives)]<=0.05)
  FN1 <- sum(padj[1:positives]>0.05)
  c(FP1,FN1)
  })
mean(result[2,]/m1)
```

